No general trend for multiple dimensions of functional diversity in bird and mammal communities

### 2024-10-14 paper/GEB Submission/manuscript_GEB.Rmd
Change hypotheses'wording

Functional diversity offers a powerful addition to species-based approaches for detecting and describing community change by capturing changes in the traits that underpin fundamental ecological processes [@lavorel2002; @mcgill2006; @suding2008].
Functional diversity metrics capture complimentary aspects of the trait space including the volume of the multidimensional space (functional diversity), how species and abundance are distributed across the space (functional evenness and divergence), and `r colorize("how the space maps to individual trait distributions", "blue")` (community weighted trait means).
Considered together, they describe the aspects of the trait space that mediate ecological process.

@@

We synthesize thousands of mammal and bird functional diversity time series to assess mean trends across communities.
We evaluate 1) whether or not there is a prevailing type of functional change across communities and if community characteristics explain variation in those trends, 2) hypotheses of change for individual trait axes, and 3) the implications of study-level trends for the ecology of those communities.
To address these goals, we present results for general trends that aggregate across all communities, category-specific trends that aggregate across communities with similar characteristics, and study-level trends.

@@

41 of the 2432 time series included in our dataset had a variable number of samples within years.
To account for this inconsistency in sampling effort we used sample-based rarefaction by bootstrap resampling within years for each time series based on the smallest number of samples in a year for that time series.
We made 200 rarefied samples for over-sampled years and calculated diversity metrics for each sample, then took the median for each metric as the value for that year and time series.
28 of the rarefied studies required under-sampling by only one sample, and only three time series required under-sampling by more than ten samples.

@@

`r colorize("Study-level trends in CWM's offer insight into how the functional space might be shifting through time, and how that functional turnover may or may not relate to species turnover. Every study that experienced a trend in a species or functional diversity metric (Table", "blue")` \ref{tab:studyChange}`r colorize("; functional change, loss of redundancy, and increase in species richness study types), with the exception of 3 \"loss of redundancy\" studies, also exhibited a trend in at least one CWM.This linking between CWM and other metrics indicates that changes in the shape of the functional space, and additions or losses of species, are almost always accompanied by a functional shift. However, shifts in the functional space were not restricted to change studies, as we also found significant CWM trends in 7 of the 32 \"no change\" studies. Functional turnover was also qualitatively divorced from species turnover, with just as many studies exhibiting functional turnover (at least one significant CWM trend) with species turnover as not. These post hoc observations give evidence that even minimal changes in species composition undetected by trends in species-based metrics can translate to significant changes to one or a few trait axes.", "blue")`

>>>>

Functional diversity offers a powerful addition to species-based approaches for detecting and describing community change by capturing changes in the traits that underpin fundamental ecological processes [@lavorel2002; @mcgill2006; @streit2023; @suding2008].
Functional diversity metrics capture complimentary aspects of the trait space including the volume of the multidimensional space (functional diversity), how species and abundance are distributed across the space (functional evenness and divergence), and `r colorize("how the space maps to individual trait distributions", "blue")` (community weighted trait means).
Considered together, they describe the aspects of the trait space that mediate ecological process.

@@

We synthesize thousands of mammal and bird functional diversity time series to assess mean trends across communities.
We evaluate 1) whether or not there is consistent change `r colorize("in the functional trait space shape (functional diversity) or location (community weighted trait means)", "blue")` across communities and if community characteristics explain variation in that change, 2) hypotheses of change for individual trait axes `r colorize("including: decreasing body size, decreasing insectivory, increasing invertivory, decreasing arboreal and understory foraging,", "blue")` and 3) the implications of study-level trends for the ecology of those communities. To address these goals, we present results for general trends that aggregate across all communities, category-specific trends that aggregate across communities with similar characteristics, and study-level trends.

@@

41 of the 2432 time series included in our dataset had a variable number of samples within years.
To account for this inconsistency in sampling effort we used sample-based rarefaction by bootstrap resampling within years for each time series based on the smallest number of samples in a year for that time series.
We made 200 rarefied samples for over-sampled years and calculated diversity metrics for each sample, then took the median for each metric as the value for that year and time series.
28 of the rarefied studies required under-sampling by only one sample, and only three time series required under-sampling by more than ten samples.
`r colorize("For time series with and without rarefaction, samples were pooled within each year and abundances were summed (where available).", "blue")`

@@

`r colorize("Study-level trends in CWM's offer insight into how the functional space might be shifting through time, and how that functional turnover may or may not relate to species turnover. Every study that experienced a trend in a species or functional diversity metric (Table", "blue")` \ref{tab:studyChange}`r colorize("; functional change, loss of redundancy, and increase in species richness study types), with the exception of 3 \"loss of redundancy\" studies, also exhibited a trend in at least one CWM. This link between CWM's and other metrics indicates that changes in the shape of the functional space, and additions or losses of species, are almost always accompanied by a functional shift. However, shifts in the functional space were not restricted to change studies, as we also found significant CWM trends in 7 of the 32 \"no change\" studies. Functional turnover was also qualitatively divorced from species turnover, with just as many studies exhibiting functional turnover (at least one significant CWM trend) with species turnover as not. These post hoc observations give evidence that even minimal changes in species composition undetected by trends in species-based metrics can translate to significant changes to one or a few trait axes.", "blue")`
